Happiness in US

### 2023-03-16 outputs/paper/Happiness-in-US-Analysis-of-the-Effects-of-Demographic-Factors-during-1972-to-2021.qmd
final paper version

---
title: "Happiness in US"
subtitle: "Analysis of the Effects of Demographic Factors during 1972 to 2021"
author: 
  - Rae Zhang
  - Faustine Fan
  - Myra Li
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

fig1 <- ggplot(happiness_data, aes(x = Year, y = People, fill = Happy_level)) +
  geom_line(aes(group = Happy_level, color = Happy_level), size = 1.2) +
  scale_fill_manual(values = happy_colors) +
  scale_color_manual(values = happy_colors) +
  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
       color = "Happy level")
fig1 <- fig1  +
  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
             fontface = "bold", 
             size = 12,
             x = 0.5,
             y = 1.05)
fig1
```
Over the past five decades, @fig-1Happiness shows there has been an overall increase in happiness levels in the United States, though with significant fluctuations. Further analysis of this trend reveals that happiness is influenced by various factors, such as macroeconomic conditions, social changes, and personal circumstances.

@@

We will examine how demographic factors such as sex, race, marital status, degree, and employment status influence happiness and why as follow.

## Sex

```{r, fig.height=6}
#| message: false
#| warning: false
#| echo: false
#| label: fig-2Gender
#| fig-cap: "Trends in happiness levels for different genders from 1972 to 2021"

@@

fig2
```
In this analysis, we examined @fig-2Gender, which compares the proportion of different education levels for each sex group from 1972 to 2021. We observed that the distribution of overall happiness levels remains consistent over time, with "pretty happy" being the most common response for both male and female groups.

@@

## Race
In @fig-3Race, we present the trends in happiness levels among different racial groups in America from 1972 to 2021. The figure shows that white people consistently report higher levels of happiness than other racial groups over the period, although this could be due to survivorship bias as the majority of respondents in the GSS dataset are white. It is important to note that all racial groups have become less happy in 2021, possibly due to the COVID-19 pandemic.

Notably, the figure shows that black people have experienced drops in happiness levels in specific years, including 1980, 2008, and 2021. These drops in happiness levels may reflect the impact of significant events or social phenomena that occurred during those years and that disproportionately affected the happiness of the black race. Further research may be needed to explore the reasons for these trends and to determine the extent to which they are influenced by factors such as racism, economic inequality, and political events.

```{r, fig.height=8}
#| message: false
#| warning: false
#| echo: false
#| label: fig-3Race
#| fig-cap: "Trends in happiness levels for different races from 1972 to 2021"

@@

fig3
```

## Degree
The trends in happiness levels across different education groups from 1972 to 2021 are presented in @fig-4Degree. It can be observed that individuals with a graduate and above education experience report the lowest levels of the "not too happy" response option and have the highest proportion of respondents selecting the "very happy" option across the years. Conversely, individuals with less than a high school education have the highest proportion of respondents selecting the "not too happy" response option among the different education groups over the years.

```{r, fig.height=16}
#| message: false
#| warning: false
#| echo: false
#| label: fig-4Degree
#| fig-cap: "Trends in happiness levels for different degrees from 1972 to 2021"

@@

fig4
```

@@

```{r, fig.height=8}
#| message: false
#| warning: false
#| echo: false
#| label: fig-5Employee
#| fig-cap: "Trends in happiness levels for different emoloyee type from 1972 to 2021"

@@

fig5
```

## Marital Status 
```{r, fig.height=16}
#| message: false
#| warning: false
#| echo: false
#| label: fig-6Marital
#| fig-cap: "Trends in happiness levels for different marital status from 1972 to 2021"

@@

fig6
```
@fig-6Marital indicates that married individuals are more likely to report higher levels of happiness, with a greater proportion of respondents selecting "Pretty happy" and "Very happy" compared to those in other marital status categories. Interestingly, the proportion of respondents selecting "Very happy" in other groups is comparatively lower. Additionally, the distribution of happiness levels in the married group appears to be more stable and consistent over time compared to the other groups.

```{r}

```

```{r}

```

@@

Research has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group (@FREDERICK20221). This can be due to factors such as experiences of racism, lack of representation in positions of power, and limited access to resources and opportunities. However, it is important to note that the impact of race identity on happiness is complex and multifaceted. The relationship between race and happiness can be influenced by factors such as social support, cultural identity, and personal values. Moreover, individual differences in personality, life circumstances, and coping strategies can also play a significant role in shaping the relationship between race identity and happiness. Therefore, while race identity can be a relevant factor in determining people's level of happiness, it is important to consider the context and individual differences in any analysis of this relationship.

## Education levels and Happiness

## Marital Status and level of Happiness

## Weakness

##Next Steps

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

\newpage
# Appendix
A link to our survey exploring what will affect happiness in US can be found in the footnotes below.^[https://forms.gle/ZAoN1uFc9Gc24anh8]

```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics(here::here("outputs/survey/survey_1.png"))
knitr::include_graphics(here::here("outputs/survey/survey_2.png"))
knitr::include_graphics(here::here("outputs/survey/survey_3.png"))
knitr::include_graphics(here::here("outputs/survey/survey_4.png"))
knitr::include_graphics(here::here("outputs/survey/survey_5.png"))
knitr::include_graphics(here::here("outputs/survey/survey_6.png"))
```
# Additional details

\newpage

# References

>>>>

---
title: "Happiness in US"
subtitle: "Analysis of the Effects of Demographic Factors during 1972 to 2021"
author: 
  - Rae Zhang
  - Faustine Fan
  - Myra Li
thanks: "Code and data are available at: https://github.com/JunweiZhang130/Happiness-in-US-Analysis-of-the-Effects-of-Demographic-Factors-during-1972-to-2021.git"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This essay explores the trends of happiness in the United States from 1972 to 2021 and identifies the significant factors that affect happiness levels. The report is based on the data gathered in US General Social Survey, and it investigates the impact of demographic features like sex, race, marital status, education, and employment types on happiness levels. The report hypothesizes that education, social isolation, and work status have a more significant impact on happiness levels than demographic features like race and gender. The COVID-19 pandemic has had overall negative effects on happiness levels. The report concludes that understanding the roots of different levels of happiness can help policymakers and society develop effective strategies to improve the quality of life for everyone."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

fig1 <- ggplot(happiness_data, aes(x = Year, y = People, fill = Happy_level)) +
  geom_line(aes(group = Happy_level, color = Happy_level), size = 1.2) +
  scale_fill_manual(values = happy_colors) +
  scale_color_manual(values = happy_colors) +
  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
       color = "Happy level") +
  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
             fontface = "bold", 
             size = 12,
             x = 0.5,
             y = 1.05)
fig1
```

Over the past five decades, @fig-1Happiness shows there has been an overall increase in happiness levels in the United States, though with significant fluctuations. Further analysis of this trend reveals that happiness is influenced by various factors, such as macroeconomic conditions, social changes, and personal circumstances.

@@

We will examine how demographic factors such as sex, race, marital status, degree, and employment status influence happiness and why as follow.
\newpage

## Sex
```{r, fig.height=6}
#| message: false
#| warning: false
#| echo: false
#| label: fig-2Gender
#| fig-cap: "Trends in happiness levels for different genders from 1972 to 2021"

@@

fig2
```

In this analysis, we examined @fig-2Gender, which compares the proportion of different education levels for each sex group from 1972 to 2021. We observed that the distribution of overall happiness levels remains consistent over time, with "pretty happy" being the most common response for both male and female groups.

@@

## Race
```{r, fig.height=8}
#| message: false
#| warning: false
#| echo: false
#| label: fig-3Race
#| fig-cap: "Trends in happiness levels for different races from 1972 to 2021"

@@

fig3
```
In @fig-3Race (next page), we present the trends in happiness levels among different racial groups in America from 1972 to 2021. The figure shows that white people consistently report higher levels of happiness than other racial groups over the period, although this could be due to survivorship bias as the majority of respondents in the GSS dataset are white. It is important to note that all racial groups have become less happy in 2021, possibly due to the COVID-19 pandemic.

Notably, the figure shows that black people have experienced drops in happiness levels in specific years, including 1980, 2008, and 2021. These drops in happiness levels may reflect the impact of significant events or social phenomena that occurred during those years and that disproportionately affected the happiness of the black race. Further research may be needed to explore the reasons for these trends and to determine the extent to which they are influenced by factors such as racism, economic inequality, and political events.
\newpage

## Degree
The trends in happiness levels across different education groups from 1972 to 2021 are presented in @fig-4Degree (next page). It can be observed that individuals with a graduate and above education experience report the lowest levels of the "not too happy" response option and have the highest proportion of respondents selecting the "very happy" option across the years. Conversely, individuals with less than a high school education have the highest proportion of respondents selecting the "not too happy" response option among the different education groups over the years.
```{r, fig.height=12}
#| message: false
#| warning: false
#| echo: false
#| label: fig-4Degree
#| fig-cap: "Trends in happiness levels for different degrees from 1972 to 2021"

@@

fig4
```
\newpage

@@

```{r, fig.height=5}
#| message: false
#| warning: false
#| echo: false
#| label: fig-5Employee
#| fig-cap: "Trends in happiness levels for different emoloyee type from 1972 to 2021"

@@

fig5
```
\newpage

## Marital Status 
@fig-6Marital (next page) indicates that married individuals are more likely to report higher levels of happiness, with a greater proportion of respondents selecting "Pretty happy" and "Very happy" compared to those in other marital status categories. Interestingly, the proportion of respondents selecting "Very happy" in other groups is comparatively lower. Additionally, the distribution of happiness levels in the married group appears to be more stable and consistent over time compared to the other groups.

```{r, fig.height=16}
#| message: false
#| warning: false
#| echo: false
#| label: fig-6Marital
#| fig-cap: "Trends in happiness levels for different marital status from 1972 to 2021"

@@

fig6
```
\newpage

@@

Research (@FREDERICK20221) has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group. This can be due to factors such as experiences of racism, lack of representation in positions of power, and limited access to resources and opportunities. However, it is important to note that the impact of race identity on happiness is complex and multifaceted. The relationship between race and happiness can be influenced by factors such as social support, cultural identity, and personal values. Moreover, individual differences in personality, life circumstances, and coping strategies can also play a significant role in shaping the relationship between race identity and happiness. Therefore, while race identity can be a relevant factor in determining people's level of happiness, it is important to consider the context and individual differences in any analysis of this relationship.

## Education levels and Happiness
Diener et al. (@Diener2013UniversalsAC) argues that education may foster a sense of personal growth, autonomy, and purpose, which can contribute to overall life satisfaction . This conclusion matches our observation of @fig-4Degree. The plot shows that higher levels of education can have a significant impact on people's level of happiness in the US. Education can provide individuals with important skills, knowledge, and opportunities that can improve their well-being and quality of life. For example, individuals with higher levels of education may have better job prospects, higher income, and greater access to resources and social support networks. These factors can contribute to higher levels of life satisfaction and happiness. One interesting fact we noticed is, less portion of people were negatively affected by the covid-19 in those who have above university degree. Which shows the skills gathered from higher education have advantages to maintain people's emotions.

Education can provide individuals with important skills, knowledge, and opportunities that can improve their well-being and quality of life (@POWDTHAVEE201510). For example, individuals with higher levels of education tend to have better job prospects, higher income, and greater access to resources and social support networks. These factors can contribute to higher levels of life satisfaction and happiness.

While our analysis suggests that education may have a positive impact on happiness levels, it is important to note that this relationship is not deterministic, and individual differences play a significant role in shaping one's happiness. As shown in @fig-3Race, even within the group of individuals with the highest level of education, there are still a significant number of respondents in every group selected "Not very happy" on the happiness scale. While education may contribute to higher levels of happiness, it is important to consider the complexity of this relationship and to recognize that happiness is a multifaceted construct that is influenced by a range of factors. This highlights the fact that education is not a guarantee of shaping one's sense of well-being.

## Marital Status and level of Happiness
@fig-6Marital illustrates the relationship between marital status and happiness levels, showing that being married or in a committed relationship is associated with higher levels of life satisfaction and happiness compared to those who are single or divorced.

There are several factors that may explain the relationship between marital status and happiness. Marital status plays a significant role in increasing happiness levels, as married people have more stable sexual partners (@MoneySexnHappiness). Marriage or a committed relationship can provide individuals with levels that may reflect the impact of significant events or social phenomena that social support, companionship, and a sense of belonging, which can contribute to feelings of happiness and well-being. Additionally, being in a stable and supportive relationship can provide individuals with a sense of security and stability that may contribute to their overall happiness.

## Weakness 
One potential weakness of this paper is that the data used may not show all the responses to the survey row by row. Additionally, the dataset only represents responses from the United States, which limits its generalizability to countries with vastly different cultural backgrounds, such as those in Asia or South America.

Furthermore, as the number of responses can vary from year to year, it's possible that the results may not be entirely objective, and errors in the dataset could occur. As a result, the conclusions drawn from this paper may not accurately reflect the overall factors that contribute to income inequality in society, and the findings may not be applicable globally.
It's important to keep in mind that the conclusions drawn from this paper are based solely on data from the United States and may not be representative of the world as a whole. Therefore, caution should be exercised when applying these findings to other countries or regions.

## Next Steps
When analyzing the relationship between marital status and happiness, it is important to consider the context and individual differences. Researchers should take into account factors such as the quality of the relationship, individual values and preferences, and other life circumstances that may impact one's level of happiness. By considering these factors, we can gain a more nuanced understanding of the relationship between marital status and happiness.
Moving forward, there are several ways to expand on this study and enhance our understanding of income inequality. First, we encourage other researchers to replicate our study using the GSS data in the Github folder to test the validity and reliability of our findings. Additionally, we included a supplementary survey in appendix 1 that contains additional data we believe would be beneficial to this study. Collecting more data can enable us to conduct more in-depth research on the factors that contribute to income inequality using the GSS data.

In addition, it would be valuable to compare our findings with those from other research papers that use national surveys in different countries to validate our conclusions and identify potential differences across regions. This could help us gain a more comprehensive understanding of the factors that contribute to income inequality on a global scale. By continuing to build on this research, we can work towards creating more equitable societies and improving the quality of life for individuals across the world.

\newpage
# Appendix
A link to our survey exploring what will affect happiness in US can be found in the footnotes below.^[https://forms.gle/ZAoN1uFc9Gc24anh8]
And you can access our survey through the QR code below.

```{r, echo=FALSE, out.width="30%"}
knitr::include_graphics(here::here("outputs/survey/QR_code.png"))
```

```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics(here::here("outputs/survey/survey_1.png"))
knitr::include_graphics(here::here("outputs/survey/survey_2.png"))
knitr::include_graphics(here::here("outputs/survey/survey_3.png"))
knitr::include_graphics(here::here("outputs/survey/survey_4.png"))
knitr::include_graphics(here::here("outputs/survey/survey_5.png"))
knitr::include_graphics(here::here("outputs/survey/survey_6.png"))
```
{{< pagebreak >}}
## References
